Type Visualization for R

### 2020-06-19 paper/main.tex
paper: Add data and collection section

@@

\section{Data}

\todo{
  Can we get a copy of Alexi's paper? This should summarize the
  dynamic analysis, what it records, and how large it is.
}

@@

Our backend is a Node.js application that serves
JSON in a standard REST-style architecture.
Data is stored in a SQLite database that has been
indexed to achieve lookups as fast as possible.

@@

% more type system features (dimensionality, dispatch)
% other languages
% general philosophy, not just about types

@@

%\section{Using the Style Template}
%
%\begin{itemize}
%\item If you receive compilation errors along the lines of ``\texttt{Package ifpdf Error: Name clash, \textbackslash ifpdf is already defined}'' then please add a new line ``\texttt{\textbackslash let\textbackslash ifpdf\textbackslash relax}'' right after the ``\texttt{\textbackslash documentclass[journal]\{vgtc\}}'' call. Note that your error is due to packages you use that define ``\texttt{\textbackslash ifpdf}'' which is obsolete (the result is that \texttt{\textbackslash ifpdf} is defined twice); these packages should be changed to use ifpdf package instead.
%\item The style uses the hyperref package, thus turns references into internal links. We thus recommend to make use of the ``\texttt{\textbackslash autoref\{reference\}}'' call (instead of ``\texttt{Figure\~{}\textbackslash ref\{reference\}}'' or similar) since ``\texttt{\textbackslash autoref\{reference\}}'' turns the entire reference into an internal link, not just the number. Examples: \autoref{fig:sample} and \autoref{tab:vis_papers}.
%\item The style automatically looks for image files with the correct extension (eps for regular \LaTeX; pdf, png, and jpg for pdf\LaTeX), in a set of given subfolders (figures/, pictures/, images/). It is thus sufficient to use ``\texttt{\textbackslash includegraphics\{CypressView\}}'' (instead of ``\texttt{\textbackslash includegraphics\{pictures/CypressView.jpg\}}'').
%\item For adding hyperlinks and DOIs to the list of references, you can use ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' (instead of ``\texttt{\textbackslash bibliographystyle\{abbrv\}}''). It uses the doi and url fields in a bib\TeX\ entry and turns the entire reference into a link, giving priority to the doi. The doi can be entered with or without the ``\texttt{http://dx.doi.org/}'' url part. See the examples in the bib\TeX\ file and the bibliography at the end of this template.\\[1em]
%\textbf{Note 1:} occasionally (for some \LaTeX\ distributions) this hyper-linked bib\TeX\ style may lead to \textbf{compilation errors} (``\texttt{pdfendlink ended up in different nesting level ...}'') if a reference entry is broken across two pages (due to a bug in hyperref). In this case make sure you have the latest version of the hyperref package (i.\,e., update your \LaTeX\ installation/packages) or, alternatively, revert back to ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-narrow\}}'' (at the expense of removing hyperlinks from the bibliography) and try ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' again after some more editing.\\[1em]
%\textbf{Note 2:} the ``\texttt{-narrow}'' versions of the bibliography style use the font ``PTSansNarrow-TLF'' for typesetting the DOIs in a compact way. This font needs to be available on your \LaTeX\ system. It is part of the \href{https://www.ctan.org/pkg/paratype}{``paratype'' package}, and many distributions (such as MikTeX) have it automatically installed. If you do not have this package yet and want to use a ``\texttt{-narrow}'' bibliography style then use your \LaTeX\ system's package installer to add it. If this is not possible you can also revert to the respective bibliography styles without the ``\texttt{-narrow}'' in the file name.\\[1em]
%DVI-based processes to compile the template apparently cannot handle the different font so, by default, the template file uses the \texttt{abbrv-doi} bibliography style but the compiled PDF shows you the effect of the \texttt{abbrv-doi-hyperref-narrow} style.
%\end{itemize}
%
%\section{Bibliography Instructions}
%
%\begin{itemize}
%\item Sort all bibliographic entries alphabetically but the last name of the first author. This \LaTeX/bib\TeX\ template takes care of this sorting automatically.
%\item Merge multiple references into one; e.\,g., use \cite{Max:1995:OMF,Kitware:2003} (not \cite{Kitware:2003}\cite{Max:1995:OMF}). Within each set of multiple references, the references should be sorted in ascending order. This \LaTeX/bib\TeX\ template takes care of both the merging and the sorting automatically.
%\item Verify all data obtained from digital libraries, even ACM's DL and IEEE Xplore  etc.\ are sometimes wrong or incomplete.
%\item Do not trust bibliographic data from other services such as Mendeley.com, Google Scholar, or similar; these are even more likely to be incorrect or incomplete.
%\item Articles in journal---items to include:
%  \begin{itemize}
%  \item author names
%	\item title
%	\item journal name
%	\item year
%	\item volume
%	\item number
%	\item month of publication as variable name (i.\,e., \{jan\} for January, etc.; month ranges using \{jan \#\{/\}\# feb\} or \{jan \#\{-{}-\}\# feb\})
%  \end{itemize}
%\item use journal names in proper style: correct: ``IEEE Transactions on Visualization and Computer Graphics'', incorrect: ``Visualization and Computer Graphics, IEEE Transactions on''
%\item Papers in proceedings---items to include:
%  \begin{itemize}
%  \item author names
%	\item title
%	\item abbreviated proceedings name: e.\,g., ``Proc.\textbackslash{} CONF\_ACRONYNM'' without the year; example: ``Proc.\textbackslash{} CHI'', ``Proc.\textbackslash{} 3DUI'', ``Proc.\textbackslash{} Eurographics'', ``Proc.\textbackslash{} EuroVis''
%	\item year
%	\item publisher
%	\item town with country of publisher (the town can be abbreviated for well-known towns such as New York or Berlin)
%  \end{itemize}
%\item article/paper title convention: refrain from using curly brackets, except for acronyms/proper names/words following dashes/question marks etc.; example:
%\begin{itemize}
%	\item paper ``Marching Cubes: A High Resolution 3D Surface Construction Algorithm''
%	\item should be entered as ``\{M\}arching \{C\}ubes: A High Resolution \{3D\} Surface Construction Algorithm'' or  ``\{M\}arching \{C\}ubes: A high resolution \{3D\} surface construction algorithm''
%	\item will be typeset as ``Marching Cubes: A high resolution 3D surface construction algorithm''
%\end{itemize}
%\item for all entries
%\begin{itemize}
%	\item DOI can be entered in the DOI field as plain DOI number or as DOI url; alternative: a url in the URL field
%	\item provide full page ranges AA-{}-BB
%\end{itemize}
%\item when citing references, do not use the reference as a sentence object; e.\,g., wrong: ``In \cite{Lorensen:1987:MCA} the authors describe \dots'', correct: ``Lorensen and Cline \cite{Lorensen:1987:MCA} describe \dots''
%\end{itemize}
%
%\section{Example Section}
%
%Lorem\marginpar{\small You can use the margins for comments while editing the submission, but please remove the marginpar comments for submission.} ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
%dolore magna aliquyam erat, sed diam
%voluptua~\cite{Kitware:2003,Max:1995:OMF}. At vero eos et accusam et
%justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
%takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
%amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
%invidunt ut labore et dolore magna aliquyam erat, sed diam
%voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est.
%
%\section{Exposition}
%
%Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
%molestie consequat, vel illum dolore eu feugiat nulla facilisis at
%vero eros et accumsan et iusto odio dignissim qui blandit praesent
%luptatum zzril delenit augue duis dolore te feugait nulla
%facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
%sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
%aliquam erat volutpat~\cite{Kindlmann:1999:SAG}.
%
%\begin{equation}
%\sum_{j=1}^{z} j = \frac{z(z+1)}{2}
%\end{equation}
%
%Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
%takimata sanctus est Lorem ipsum dolor sit amet.
%
%\subsection{Lorem ipsum}
%
%Lorem ipsum dolor sit amet (see \autoref{tab:vis_papers}), consetetur sadipscing elitr, sed diam
%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
%takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
%amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
%invidunt ut labore et dolore magna aliquyam erat, sed diam
%voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum.
%
%\subsection{Mezcal Head}
%
%Lorem ipsum dolor sit amet (see \autoref{fig:sample}), consetetur sadipscing elitr, sed diam
%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
%takimata sanctus est Lorem ipsum dolor sit amet.
%
%\subsubsection{Duis Autem}
%
%Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
%takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
%amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
%invidunt ut labore et dolore magna aliquyam erat, sed diam
%voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est. Lorem
%ipsum dolor sit amet.
%
%\begin{figure}[tb]
% \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
% \includegraphics[width=\columnwidth]{paper-count-w-2015-new}
% \caption{A visualization of the 1990--2015 data from \autoref{tab:vis_papers}. The image is from \cite{Isenberg:2017:VMC} and is in the public domain.}
% \label{fig:sample}
%\end{figure}
%
%\subsubsection{Ejector Seat Reservation}
%
%Duis autem~\cite{Lorensen:1987:MCA}\footnote{The algorithm behind
%Marching Cubes \cite{Lorensen:1987:MCA} had already been
%described by Wyvill et al. \cite{Wyvill:1986:DSS} a year
%earlier.} vel eum iriure dolor in hendrerit
%in vulputate velit esse molestie consequat,\footnote{Footnotes
%appear at the bottom of the column.} vel illum dolore eu
%feugiat nulla facilisis at vero eros et accumsan et iusto odio
%dignissim qui blandit praesent luptatum zzril delenit augue duis
%dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet,
%consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt
%ut laoreet dolore magna aliquam erat volutpat.
%
%
%\paragraph{Confirmed Ejector Seat Reservation}
%
%Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
%suscipit lobortis nisl ut aliquip ex ea commodo
%consequat~\cite{Nielson:1991:TAD}. Duis autem vel eum iriure dolor in
%hendrerit in vulputate velit esse molestie consequat, vel illum dolore
%eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
%dignissim qui blandit praesent luptatum zzril delenit augue duis
%dolore te feugait nulla facilisi.
%
%\paragraph{Rejected Ejector Seat Reservation}
%
%Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
%suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
%vel eum iriure dolor in hendrerit in vulputate velit esse molestie
%
%
%\section{Conclusion}
%
%Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
%takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
%amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
%invidunt ut labore et dolore magna aliquyam erat, sed diam
%voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum.

%% if specified like this the section will be committed in review mode
\acknowledgments{
The authors wish to thank A, B, and C. This work was supported in part by
a grant from XYZ.}

%\bibliographystyle{abbrv}
\bibliographystyle{abbrv-doi}
%\bibliographystyle{abbrv-doi-narrow}
%\bibliographystyle{abbrv-doi-hyperref}
%\bibliographystyle{abbrv-doi-hyperref-narrow}

>>>>

% R background

% gradual typing background

% vis papers

@@

\section{Data and Collection}

\todo{Alexi's paper says 412, but I see more packages in our data.}

The data consists of traces from 412 packages
containing over $760,000$ lines of R code
and $534,000$ lines of native code.
All libraries were obtained from CRAN,
the most popular R code repository.
Packages hosted on CRAN must satisfy certain criteria
to ensure they are high quality.
Additionally, the subset of 412 packages chosen here
all had code coverage above 65\%.

Each execution trace was recorded by running
the test, example, and vignette code of each package.
A vignette is a form of documentation that weaves
together prose and executable code.
An execution trace is a list of ``raw'' type signatures
containing, among other attributes, the function name
and type tags for each argument and return value.
Our dataset consists of a table of traces
that have been processed to reduce some fields.
Traces were generated via a dynamic analysis
that instruments the R virtual machine with hooks
that intercept the required data at run-time.

While {\sc TypeVis} has only been run with
data from R execution traces, the architecture is not
limited to R.
Any dynamically typed language with the right run-time
hooks could generate data for the tool.
As long as execution trace information is adheres to
the same database schema, {\sc TypeVis} will
work properly without any modifications.

@@

Our backend is a Node.js application that serves
JSON in a standard REST-style architecture.
Data is stored in a SQLite database that has been
indexed to achieve lookups as fast as possible.
To avoid execessive memory consumption and speed up computations,
clients query data from the server on an as-needed basis.
Thus, a stable network connection is required for
real-time interaction.

@@

% other dynamic languages

% other featuers not just types

% general philosophy not just about types
% vis has role to play in language design

@@

\acknowledgments{
  The authors wish to thank A, B, and C. This work was supported in part by
  a grant from XYZ.
}

\bibliographystyle{abbrv-doi}
